High-throughput Genomic Paper Notes

### 2017-11-19 Paper_notes/Paper_notes.Rmd
Update Paper_notes.Rmd

---
title: "High-throughput Genomic Paper Notes"
author: "Jonathan Yu"
date: "October 25, 2017"
output: pdf_document
---

@@

@@

- Bolstad, B. M., R. A. Irizarry, M. Astrand, and T. P. Speed. “A Comparison of Normalization Methods for High Density Oligonucleotide Array Data Based on Variance and Bias.” Bioinformatics (Oxford, England) 19, no. 2 (January 22, 2003): 185–93. - Normalization methods description.

@@

- Cleveland, William S, and Susan J Devlin. “Locally Weighted Regression: An Approach to Regression Analysis by Local Fitting.” Journal of the American Statistical Association 83, no. 403 (1988): 596–610. - Loess regression. Concept, statistics.

>>>>

---
title: "High-throughput Genomic Paper Notes"
author: "Jonathan Yu"
date: "November 19, 2017"
output: pdf_document
---

@@

### Smyth, 2004

- Smyth, Gordon K. "Linear Models and Empirical Bayes Methods for Assessing Differential Expression in Microarray Experiments." Statistical Applications in Genetics and Molecular Biology 3 (2004)

Smyth proposed an extended hierarchical model for microarray experiments. Given the large number of gene-wise linear model fits, a moderated t, using the classical t-statistic as a basis, uses a posterior variance instead of the sample variance to detect differences in gene expressions. The innovation is that prior information is assumed for the variance among the genes and thus, the posterior can be derived to shrink the observed variances. The hypothesis would be that model coefficient that distinguishes differences between group sis zero. It was noted that when gene identifcation using either the current B-statistic model or the moderated t-statistic, they would lead to similar results albeit the t-statistic method would require less knowledge of the parameters. Simulation shows that a moderated t has a lower false discovery rate than other methods and that it can be used for single and two color microarray experiments.

### Tusher, Tibshirani, and Chu, 2001

- Tusher, V. G., R. Tibshirani, and G. Chu. "Significance Analysis of Microarrays Applied to the Ionizing Radiation Response." Proceedings of the National Academy of Sciences of the United States of America 98, no. 9 (April 24, 2001)

the method Significance Analysis of Microarrays (SAM) was developed to identify significant changes in the expression of thousand of genes during the many different biological states. A score is assigned to each gene for each change relative to the standard deviation for repeated measurements. The genes with scores above a specified threshold, permutations of the repeated measurements are used to estimate the false discovery rate (FDR).

First, the score comes from the signal-to-noise ratio where they ccounted for gene-specific fluctionations in the gene expressions d(i) $d(i) = \frac{\bar{x}_I(i) - \bar{x}_U(i)}{s(i) + s_0}$ where the numerator is the difference between the average levels of gene i between two states I and U. The gene variation of the repeated measurement account is used in the denominator. Next, the differences are ranked from highest different to lowest relative differnece. Then, permutations of the repeated measurements were done in order to minimize potential confounding effects from differences between two cell lines. A statistic is calculated for each permutation which will result in a distribution of the relative differences where the mean reltaive difference can be derived. This can then compare and identify induced and repressed genes. Their method "for setting thresholds provides asymmetric cutoffs for induced and repressed genes.". The alternative method, t test, has a symmetric horizontal cutoff. SAM has neither strong or weak family wise error rate.

@@

- Bolstad, B. M., R. A. Irizarry, M. Astrand, and T. P. Speed. "A Comparison of Normalization Methods for High Density Oligonucleotide Array Data Based on Variance and Bias." Bioinformatics (Oxford, England) 19, no. 2 (January 22, 2003): 185-93. - Normalization methods description.

@@

- Cleveland, William S, and Susan J Devlin. "Locally Weighted Regression: An Approach to Regression Analysis by Local Fitting." Journal of the American Statistical Association 83, no. 403 (1988): 596-610. - Loess regression. Concept, statistics.
